---
title: "Hypothesis Test 1 for Employee Data Set"
output: word_document
---

```{r}
library(ggplot2)
library(ggpubr)
library(dplyr)
```
The question we wish to answer with our first hypothesis is - is there a significant difference between men and women when it comes to a yearly PercentageSalaryHike (PSH).

This is a non trivial problem for any company as a gender wage gap would lead to loss in talent.

```{r}

#We first take out dataset and split it into data for both men and women
attrition_data <- read.csv('C:\\Users\\bcava\\Desktop\\AMS572 Project\\WA_Fn-UseC_-HR-Employee-Attrition.csv')
names(attrition_data)[1] <- "Age"
#View(attrition_data)
```



First, let us do our Exploratory Data Analysis (EDA)

```{r}
TAB <- table(attrition_data$Gender, attrition_data$PercentSalaryHike)
TAB
```

```{r}

barplot(TAB, beside = T, legend = T)
```


We are NOW going to do a hypothesis test for two sample means - We will take the mean of the PercentSalaryHike (PSH) for both men and women and see if there is any significant difference

If we consider the mean of men's PSH to be u1 and the mean of women's PSH to be u2 then our hypothesis is 
H0 : u1 equals u2 (u1-u2 = 0)
H1 : u1 not equal to u2 (u1- u2 != 0)


We split our data into male and female 


```{r}
X<-split(attrition_data, attrition_data$Gender)
#View(X$Female)
#View(X$Male)

```

Firstly, we need to check for normality on our data. We will conduct a Shapiro Wilks test on our data to see if we can consider this data to have a normal distribution
```{r}
shapiro.test(X$Male$PercentSalaryHike)
shapiro.test(X$Female$PercentSalaryHike)
summary(X$Male$PercentSalaryHike)
summary(X$Female$PercentSalaryHike)
```

As we can see, the p value is much lesser than 0.05 so this does not follow a normal distribution

Test of Normality II - visual
```{r}
library(ggplot2)
library(ggpubr)
ggdensity(X$Male$PercentSalaryHike, main = "Normality", xlab = "PSH for men")
ggdensity(X$Female$PercentSalaryHike, main = "Normality", xlab = "PSH for women")
```

Since our data is so large and is still not normally distributed, we will follow a Wilcoxon Rank-Sum Test (Or Man Whiteney Test) to test our hypothesis

```{r}
summary(X$Male$PercentSalaryHike)
summary(X$Female$PercentSalaryHike)
```
Visualizing our data with boxplots
```{r}
ggboxplot(
  attrition_data,
  x= "Gender",
  y = "PercentSalaryHike",
  combine = FALSE,
  merge = FALSE,
  color = "black",
  fill = "#E7B800",
  palette = NULL,
  title = "BoxPlot Comparison of PSH",
  xlab = "Gender",
  ylab = "PSH")
```




```{r}
wilcox.test(X$Male$PercentSalaryHike,X$Female$PercentSalaryHike ) 
```
The p-value here is 0.7043 which is much higher than our alpha value of 0.1 for a 10% significance level so we cannot reject the null hypothesis 

Hence we cannot conclude that there is any major difference between men and women when it comes to a percentage salary hike - This is important as if we just look at our EDA, we may assume that there is a significant difference

---