---
title: "R Notebook"
output: word_document
---


```{r setup, include= FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r}
library(plyr)
library(data.table)
library(dplyr)
library(VIM)
library(DT)
library(gridExtra)
library(ggplot2)
library(caret)
library(Metrics)
library(randomForest)
library(pROC)
library(e1071)
library(dtree)
library(corrplot)
library(Amelia)
library(caTools)
library(caret)
library(car)
library(mice)
library(missMethods)
knitr::opts_chunk$set(warning = FALSE)
knitr::opts_chunk$set(message = FALSE)
```

```{r}
attrition_data <- read.csv('C:\\Users\\bcava\\Desktop\\AMS572 Project\\WA_Fn-UseC_-HR-Employee-Attrition.csv')
names(attrition_data)[1] <- "Age"
```

- Variables type:
  - **Numeric variables**: 
    - Related to personal information: age, distance_from_home, employee_number (id variable)
    - Related to income: hourly_rate, daily_rate, monthly_rate, monthly_income, percent_salary_hike
    - Related to time in company: years_at_company, years_in_current_role, years_since_last_promotion, years_with_curr_manager, total_working_years
    - other: num_companies_worked, standard_hours(to delete), training_times_last_year, employee_count (to delete)
  - **Categorical variables**: 
    - **Binary variables**: attrition(target variable),gender, over18 (to delete),over_time
    - **Nominal variables**: department, education_field, job_role, marital_status
    - **Ordinal variables**: 
      - Ordinal regarding satisfaction and performance : environment_satisfaction, job_satisfaction, relationship_satisfaction, work_life_balance,job_involvement,performance_rating
      - Other ordinal: business_travel, education, job_level, stock_option_level



#checking for the null values
```{r}
apply(is.na(attrition_data), 2, sum)
```
No null values in our data

# CheckING for duplicated Record
```{r}
sum (is.na(duplicated(attrition_data)))
```


### Converting Categerical data into factors
```{r}
attrition_data$BusinessTravel = as.factor(attrition_data$BusinessTravel)
attrition_data$Department = as.factor(attrition_data$Department)
attrition_data$Gender = as.factor(attrition_data$Gender)
attrition_data$JobRole = as.factor(attrition_data$JobRole)
attrition_data$MaritalStatus = as.factor(attrition_data$MaritalStatus)
attrition_data$OverTime = as.factor(attrition_data$OverTime)
attrition_data$EducationField = as.factor(attrition_data$EducationField)
attrition_data$Attrition = as.factor(attrition_data$Attrition)
str(attrition_data)

```

```{r}
library(mice)

```

```{r}

dwm <- attrition_data #Created a copy so imputation can be compared with the original

TAB <- table(dwm$Gender, dwm$PercentSalaryHike)
TAB
```


```{r}

#Removing the 20% of the data who has the PercentSalaryHike as 11

dwm <- delete_MNAR_censoring(dwm, 0.2,"PercentSalaryHike")
```


```{r}
sapply(dwm, function(x) sum(is.na(x)))
```




```{r}

init = mice(dwm, maxit=0) 
meth = init$method
predM = init$predictorMatrix
```

```{r}

#Variables to predict
meth[c("PercentSalaryHike")]="norm" 
#View(dwm)

```


```{r}
set.seed(45)
imputed = mice(dwm, predictorMatrix=predM, m=40, seed=10, meth = 'pmm')
imputed$meth
```

```{r}
#View(dwm)
imputed <- complete(imputed)
sapply(imputed, function(x) sum(is.na(x)))
```


```{r}
X<-split(imputed, imputed$Gender)
```

```{r}
wilcox.test(X$Male$PercentSalaryHike,X$Female$PercentSalaryHike ) 
summary(X$Male$PercentSalaryHike)
summary(X$Female$PercentSalaryHike)
```


